\documentclass[DIV=8, parskip=half, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother
\newcommand*{\meta}[1]{\textlangle\textsl{#1}\textrangle}
\newcommand*{\marg}[1]{\texttt{\{}\meta{#1}\texttt{\}}}

\addtokomafont{title}{\rmfamily}

\title{The \pkg{jneurosci} package\thanks{This manual corresponds to \pkg{jneurosci}~v1.00, dated~2004 April 06.}}
\author{Matthias Hennig (\mail{hennig@cn.stir.ac.uk})}
\date{2004 April 06}


\begin{document}

\maketitle

\noindent
This is a slightly modified version of the \pkg{namedplus} style, which fully conforms
with the Journal of Neuroscience citation style.

The following citation labels are implemented:
%
\begin{labeling}{\cmd{\citeauthoryear}\marg{author-info}\marg{year}:}
\item[\cmd{\citeauthoryear}\marg{author-info}\marg{year}:] these labels are processed by the following commands: 
\item[\cmd{\cite}\marg{key}:] which produces citations with both author and year, enclosed in parens. 
\item[\cmd{\shortcite}\marg{key}:] which produces citations with year only, enclosed in parens 
\item[\cmd{\citeauthor}\marg{key}:] which produces the author information only 
\item[\cmd{\citeyear}\marg{key}:] which produces the year information only 
\item[\cmd{\citetext}\marg{key}:] which produces Author (Year) 
\item[\cmd{\citenoparens}\marg{key}:] which produces Author, Year 
\end{labeling}

The output looks like that:
%
\begin{labeling}{\cmd{\citetext}\marg{key}:}
\item[\cmd{\cite}\marg{key}:] (Strettoi and Masland, 1996; Kolb, 1997; Masland, 2001) and (Fisher et al., 1975)
\item[\cmd{\citetext}\marg{key}:] for reviews, see McNaughton (1990); Müller and Kaupp (1998); Fain et al. (2001).
\end{labeling}

The bibliography follows J~Neurosci conventions, e.\,g.:
%
\begin{itemize}
\item Koch C (1999) Biophysics of Computation: Information Processing in Single Neurons Oxford University Press.
\item Enroth-Cugell C, Robson JG (1966) The contrast sensitivity of retinal ganglion cells of the cat. J~Physiol 187:517-552. 
\end{itemize}
 
This is a modified version of the \pkg{namedplus} style by:

\begin{tabular}{@{}ll@{}}
  \toprule
  A. David Redish & \mail{adr@nsma.arizona.edu}                               \\
  Post-doc        & \url{http://www.cs.cmu.edu/~dredish}                      \\
  \multicolumn{2}{@{}l@{}}{Neural Systems, Memory and Aging, Univ of AZ, Tucson AZ} \\
  \bottomrule
\end{tabular}

which was in turn modified by:\\
Oren Patashnik (\mail{PATASHNIK@SCORE.STANFORD.EDU})

and finally modified by:\\
Matthias H Hennig, University of Edinburgh, \mail{mhennig@inf.ed.ac.uk}

\end{document}

