\documentclass[pagesize=auto, fontsize=12pt, DIV=11]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother
\newcommand*{\meta}[1]{\textlangle\textsl{#1}\textrangle}
\newcommand*{\marg}[1]{\texttt{\{}\meta{#1}\texttt{\}}}

\addtokomafont{title}{\rmfamily}

\title{The \pkg{needspace} package\thanks{This manual corresponds to \pkg{needspace}~v1.3b, dated~2009/09/03.}}
\author{Author: Peter Wilson, Herries Press\\Maintainer: Will Robertson\thanks{\mail{will.robertson@latex-project.org}}}
\date{2009/09/03}


\begin{document}

\maketitle

\noindent
Provides commands to reserve space at the bottom of a page. If there
is not enough space on the current page (column) a new page (column)
is started.

\begin{quote}
  \small
  Copyright 1996--2003 Peter Wilson\\
  Released under the \LaTeX\ Project Public License
\end{quote}


\section{Example uses:}

\begin{verbatim}
... previous text.\par
\needspace{5\baselineskip} % need next 5 lines on a single page/column
The next five lines ...

\Needspace*{4\baselineskip} % need 4 lines kept on same page
   The major principles are:
\begin{itemize}
\item ...
\end{verbatim}

The \cmd{\needspace} macro reserves an \emph{approximate} amount of space (depending
on what penalties are in effect) but is efficient and normally adequate.
A short page will be ragged bottom even if \cmd{\flushbottom} is in
effect.

\cmd{\Needspace} is less efficient but reserves the requested space. It
should only be used between paragraphs. Short pages produced by 
\cmd{\Needspace}\marg{length} will be ragged bottom but short pages produced by 
\cmd{\Needspace*}\marg{length} will be flush bottom if \cmd{\flushbottom} is in effect.


\section{Changes:}

\begin{description}
\item[Changes v1.3b (2009/09/03)]
  New maintainer (Will Robertson)
  
\item[Changes v1.3a (2003/02/18)]
  At the suggestion of Donald Arseneau (private emails) changed guts of
  \cmd{\needspace} from 
  \ldots\verb+\vskip\z@\@plus\dimen@\pagebreak[3]+\allowbreak
  \verb+\vskip\z@\@plus-\dimen@\endgroup}+

\item[Changes v1.3 (2003/02/07)]
  Added \cmd{\Needspace} (an extended version of postings by Donald Arseneau 
  to \texttt{comp.text.tex} e.\,g.,\ in November 1993, August 1995, September 1997, \ldots)
  
\item[Changes v1.2 (2002/09/11)]
  At the suggestion of Robin Fairbairns, added use of \cmd{\dimen@} to capture 
  the argument so that \pkg{calc} package commands can be used in the argument.
\end{description}

\end{document}
