\documentclass[pagesize=auto, fontsize=12pt, headings=normal]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{array}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\bmail}[1]{\href{mailto:#1}{\texttt{<#1>}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother

\addtokomafont{title}{\rmfamily}

\title{The \pkg{r\_und\_s} package\thanks{This manual corresponds to \pkg{r\_und\_s.sty}~v1.3i, dated~September 1, 2005.}}
\subtitle{R- and S-phrases for \LaTeX}
\author{Thiemo Nordenholz\thanks{\mail{nz@thiemo.net}}}
\date{September 1, 2005}


\begin{document}

\maketitle
\tableofcontents


\section{Copyright and disclaimer}

\begin{quote}
  \footnotesize
  Copyright © 1997–2005 Thiemo Nordenholz \bmail{nz@thiemo.net} \\
  All rights reserved.

  Redistribution and use in source and binary forms, with or without
  modification, are permitted provided that the following conditions
  are met:
  %
  \begin{itemize}
  \item Redistributions of source code must retain the above copyright
    notice, this list of conditions and the following disclaimer.
  \item Redistributions in binary form must reproduce the above copyright
    notice, this list of conditions and the following disclaimer in the
    documentation and/or other materials provided with the distribution.
  \end{itemize}

  \SetTracking[spacing={700*,400*,}]{encoding=*, family=*}{60}%
  \lsstyle
  THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
  ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
  ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
  FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
  DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
  OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
  HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
  LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
  OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
  SUCH DAMAGE.
\end{quote}


\section{Description}

This packet provides \LaTeX- resp.\ \TeX-Commands, which will expand numeric R-
and S-codes to a full text form. This is usable for creating safety data
sheets and lab protocols. The \texttt{*.sty} files are for \LaTeX, the \texttt{*.tex} files are
plain~\TeX.


\subsection{Languages}

The following languages are supported and contained in the mentioned files:

\medskip
\begin{tabular}{@{}l>{\ttfamily}l@{}}
  German  & r\_und\_s \\
  English & eng\_rs   \\
  Dutch   & nl\_rs    \\
  French  & fr\_rs    \\
\end{tabular}


\section{Usage}

For information on how to use external macros in \LaTeX\ and \TeX, please
consult one of the many documentations written on this subject. 
The phrases S\,5, S\,6, S\,14, S\,28, S\,40, S\,43, S\,47, S\,48, S\,50, and the combinations
S\,3/14, S\,3/9/14, S\,3/9/14/19, and S\,47/49 need a parameter specified by the
manufacturer of the product.


\section{Coding of the numbers}

First, the phrases contained in this package are divided into \enquote{R-} and \enquote{S-}
prases, beginning \enquote{\cmd{\cr}} and \enquote{\cmd{\cs}}. The \enquote{c} at the beginning stands for
\enquote{Chemie} or \enquote{chemistry} to avoid conflicts with other \LaTeX-commands.

Because the numerical coding of the R- and S-phrases is not possible with
\LaTeX, the phrases are \enquote{numbered} with letters (unfortunately, the coding
requires the knowledge of a little bit germany, but it's really easy\dots)

\medskip
\begin{tabular}{@{}l@{$\;\to\;$}>{\ttfamily}ll@{}}
  0 & o & (the \enquote{o} looks like a zero, doesn't it?)                               \\
  1 & e & (\enquote{eins} is german for \enquote{one)}                                   \\
  2 & z & (\enquote{zwei}, what means \enquote{two})                                     \\
  3 & d & (\enquote{drei})                                                               \\
  4 & v & (\enquote{vier})                                                               \\
  5 & f & (\enquote{fünf})                                                               \\
  6 & s & (\enquote{sechs})                                                              \\
  7 & i & (\enquote{sieben}, but \enquote{s} is already used, so took the second letter) \\
  8 & a & (\enquote{acht});                                                              \\
  9 & n & (\enquote{neun} and, how nice, \enquote{nine} :-)
\end{tabular}

\medskip
\noindent
For example, R-phrase~2 becomes \cmd{\crz}, R-phrase~17 becomes \cmd{\crei}, which reads
\enquote{chemistry, r-phrase eins-sieben (one-seven)}.

Combined phrases need the additional character \enquote{c} (for \enquote{combined}), e.\,g.\ %
R\,23/24/25 becomes \enquote{\cmd{\crczdzvzf}} (chemistry, r-phrase, combined, zwei-drei-
zwei-vier-zwei-fünf (two-three-two-four-two-five).

Seems difficult? Maybe, but I've gotten used to it, and so will you ;-)


\section{Completeness}

The package contains R- and S-phrases up to R\,64 and S\,62 and their most
important combinations.
In the French and Dutch version, the combination S\,37/38/39 is missing and
included in English only, marked with \enquote{XXX}.


\section{Comments}

Please send any comments to my email address given below, I will take a look
at them and promise to send at least a short confirmation when I recieved the 
mail.

\begin{quote}
  \begin{tabbing}
    Thiemo Nordenholz, \= Clausthal-Zellerfeld, den 03. März 1999 \\
    updated:           \> Hamburg, den 01. Dezember 2005
  \end{tabbing}
  %
  eMail: \mail{nz@thiemo.net}

  translated into \TeX-macro by \mail{eike.kunst@tu-clausthal.de}
\end{quote}


\end{document}
