\documentclass[pagesize=auto, fontsize=12pt, DIV=10]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{hologo}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother

\addtokomafont{title}{\rmfamily}

\title{The \pkg{concrete} package\thanks{This manual corresponds to \pkg{concrete.sty}, dated~1996/10/08.}}
\author{Jim Hefferon\thanks{\mail{jhefferon@smcvt.edu}}}
\date{1996/10/08}


\begin{document}

\maketitle

\noindent
The \pkg{concrete} package provides fonts for an approximation of the style in \emph{Concrete Mathematics.}
It simply loads the packages \pkg{beton} and \pkg{euler}; see the manuals of those two packages for details.


\section{History}

\begin{description}
\item[96-10-08 Jim Hefferon \mail{jhefferon@smcvt.edu}] \leavevmode \\
  It is an FAQ on \texttt{comp.text.tex} that people look on CTAN for a style 
  with this name.  Seems reasonable.  They are always advised to 
  \cmd{\usepackage} the two packages \pkg{beton} and \pkg{euler}.  So here is a wrapper style
  to do that.
\end{description}


\section{Remarks}

\begin{description}
\item[96-10-08 jh] \leavevmode \\
  There are other styles for Concrete Math (\pkg{concmath} and something
  from \hologo{AmSLaTeX}\,1.1\ are the ones that I know of), but this one is
  modular in that if the packages it calls are updated then of
  course, this is updated also).  In particular, there are no font 
  parameters, etc.
\end{description}


\section{Bugs}

\begin{itemize}
\item Unlike the \hologo{AmSLaTeX}\,v.\,1.1\ style file, this makes no attempt to
  provide any grafitti, etc.\ macros.  Just the fonts.  (One could
  argue that those are document-style, not font, macros.)
\end{itemize}

\end{document}
