\documentclass[DIV=9, parskip=half, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother
\addtokomafont{title}{\rmfamily}

\title{The \pkg{magaz} package\thanks{This manual corresponds to \pkg{magaz}~v0.2, dated~04--Sep--2003.}}
\author{Donald Arseneau,   \mail{asnd@triumf.ca},  Vancouver, Canada}
\date{04--Sep--2003}


\begin{document}

\maketitle

\begin{quote}
  \small
  This software is free to distribute, unaltered, and to use in any 
  application. It may be altered and incorporated in other software freely.
\end{quote}

This version is just a single tool from a larger body of tools I might
try to implement if I ever have the spare time.  I'm writing this part
now [1999] in response to a request on \texttt{comp.text.tex}.

\begin{tabular}{@{}rl@{}}
  Yes & Special formatting for first line of text in a paragraph          \\
  No  & Special formatting for for first $n$ lines (allow each different) \\
  No  & Also for first character (drop caps) (working together)           \\
  No  & ``Continued on''/``continued from'' markers                       \\
  No  & wrapfig spanning columns                                          \\
  No  & wrapfig placed at fixed location on page                          \\
  NO! & Piecing together fragments of gallies to fit page
\end{tabular}


\minisec{Instructions:}

The ``\cmd{\FirstLine}'' command takes one argument: Some text.

The first line of this text will be formatted according to
``\cmd{\FirstLineFont}''; but if the text occupies less than a line, all 
that text (and no more) will be affected by ``\cmd{\FirstLineFont}''.  
The argument should be ordinary text, not lists and displayed 
equations. Such things will give you a `Bad text' error.

This command is intended to be used in a replacement for a normal 
sectioning command in a document class; ``\cmd{\paragraph}'' would be 
appropriate.  In particular, this ``\cmd{\paragraph}'' (or whatever) should 
insert vertical space and penalties etc.,\ and probably ``\cmd{\noindent}''
at the start of the argument.  The document class should also define 
``\cmd{\FirstLineFont}'' appropriately as either a switch or a text-command; 
as in ``\verb+\renewcommand\FirstLineFont{\MakeUppercase}+''.  The defult 
definition is ``\verb+\providecommand\FirstLineFont{\scshape}+''.

\end{document}
