\documentclass[DIV=9, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\AkkTeX}{Akk\kern-0.09em\TeX\@\xspace}

\pdfstringdefDisableCommands{%
  \def\AkkTeX{AkkTeX\xspace}%
}

\title{\AkkTeX\ -- \LaTeX\ utilities by Igor Akkerman (\url{http://www.akktex.org/})}
\author{Igor Akkerman}
\date{2005/05/05}


\begin{document}

\maketitle

\section{Main Features:}

\begin{itemize}
\item new document classes for technical documents, thesis works, mathematical
  manuscripts and lecture notes
\item many math packages providing a tremendous nuber of macros for mathematical
  texts
\item layout providing a non-empty parskip with extended length corrections and
  new section definition commands
\item easy label creation for counters
\item German language tools and predefined abbreviations
\end{itemize}


\section{\AkkTeX summary}

\AkkTeX consists of a collection of new \LaTeX\ document classes and a very large
set of \LaTeX\ packages with hundreds of new macros simplifying the creation of
scientific papers. The packages have all been created within years of practical
work with \LaTeX.

\AkkTeX\ provides document classes for mathematical lecture notes, elegant
working sheets, thesis works and many more. They are designed for easy-to-use
generation of the specified kinds of documents.

Additionally, the \AkkTeX\ packages provide a tremendous number of macros,
especially for mathematic texts. Their goal is to reduce the creation time of
documents to a minimum, such that, for example, to type a mathematic paper does
not take longer than to write it by hand. Another advantage over usual \LaTeX\ %
typing is that the macro names are subject-specific, not symbol-oriented.
That is, the author can think less about the \LaTeX\ symbol names, but instead,
can call them by their meaning. The \verb+\set+ macro is just the simplest example.
Not only typing becomes more intuitive, but also, reading the \LaTeX\ source code,
changing and refactoring it is made easier.

The latest release can be found at:
\url{http://www.akktex.org/}
     
Browse our CVS repository for the latest versions of the packages and document
classes at:
\url{http://sourceforge.net/cvs/?group_id=160672}

\bigskip
\noindent\null\hfill$*\quad*\quad*\quad$\hfill\null\par
\bigskip

\noindent
This work may be distributed and/or modified under the
conditions of the \LaTeX\ Project Public License (latest version).
The latest version of this license can be found at
\url{http://www.latex-project.org/lppl.txt}

This file may only be distributed together with a copy of the
`\AkkTeX' bundle.
The files belonging to the `\AkkTeX' bundle are listed above in this file.

\end{document}
