\documentclass[DIV=8, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage[unicode=true]{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother
\newcommand*{\env}[1]{\texttt{#1}}
\newcommand*{\opt}[1]{\texttt{#1}}
\newcommand*{\meta}[1]{\textlangle\textsl{#1}\textrangle}
\newcommand*{\marg}[1]{\texttt{\{}\meta{#1}\texttt{\}}}

\addtokomafont{title}{\rmfamily}

\title{The \pkg{cuted} package}
\author{Sigitas Tolušis\\\mail{sigitas@vtex.lt}\\\url{http://www.vtex.lt/tex/download/macros/}}
\date{1997/10/01}


\begin{document}

\maketitle

\section{Purpose:}

Insert some onecolumn material in full-width on double-column page.


\section{Short Description:}

It is denoted to place some material in full-width at any place 
on double-column page (\cmd{\twocolumn} mode). The part of material
can be placed on the next page. Existing text is placed
under inserted material in two columns (it is reformatted).

Provides new environment:
\verb+\begin{strip}+\meta{\ldots}\verb+\end{strip}+\\
Puts \cmd{\vbox}\marg{\ldots} in full-width on double-column page.
Adds \cmd{\stripsep} glue before and after
(By default \verb+\stripsep=15pt plus2pt minus 2pt+).
It can be problems with floats and footnotes. In the case of 
problems with footnotes it is possible to correct the situation
using commands \cmd{\footnotemark} and \cmd{\footnotetext}.

Provides new command:
\cmd{\oldcolsbreak}\marg{\#1}\\
Adds \meta{\#1} in place of original column break if strip is placed 
on the right column.


\section{List of Changes}

\begin{tabular}{@{}l@{ -- }l@{}}
  1997/09/02 & improved column balancing                        \\
  1997/09/03 & float control for \cmd{\@colroom} setting        \\
  1997/09/04 & improved \cmd{\@colroom} setting                 \\
  1997/09/05 & checking for \cmd{\@viper} and \cmd{\hold@viper} \\
  1997/09/09 & compatibility with \pkg{flushend.sty}            \\
  1997/09/22 & improved \cmd{\output} ir \cmd{\@specialoutput}  \\
  1997/10/01 & improved \cmd{\output} ir \cmd{\@addviper}
\end{tabular}

\bigskip

\footnotesize
\noindent
\pkg{cuted.sty}\\
Copyright 1997 Sigitas Tolušis\\
V\TeX\ Ltd., Akademijos 4, Vilnius, Lithuania\\
e-mail \mail{sigitas@vtex.lt}\\
\url{http://www.vtex.lt/tex/download/macros/}

This program can redistributed and/or modified under the terms
of the \LaTeX\ Project Public License Distributed from CTAN
archives in directory \texttt{macros/latex/base/lppl.txt}; either
version~1 of the License, or (at your option) any later version.

\end{document}


